Fast Dictionary Learning for High-Dimensional Seismic Reconstruction

### 2020-11-09 sgk5d/paper.tex
2020年11月 8日 星期日 17时47分27秒 CST
2020年11月 9日 星期一 11时05分45秒 CST

\title{Fast dictionary learning for high-dimensional seismic reconstruction}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
%\author{Hang Wang, Quan Zhang, Wei Chen, Xingye Liu, Shaohuan Zu, and Yangkang Chen
\author{Hang Wang, Wei Chen, Quan Zhang, Xingye Liu, Shaohuan Zu, and Yangkang Chen
\thanks{H. Wang, Q. Zhang, and Y. Chen are with the Key Laboratory of Geoscience Big Data and Deep Resource of Zhejiang Province, School of Earth Sciences, Zhejiang University (chenyk2016@gmail.com).}
\thanks{W. Chen is with the Key Laboratory of Exploration Technology for Oil and Gas Resources, Ministry of Education, Yangtze University, and also with the Hubei Cooperative Innovation Center of Unconventional Oil and Gas, Yangtze University.}
\thanks{X. Liu is with College of Geology and Environment, Shaanxi Provincial Key Laboratory of Geological Support for Coal Green Exploitation, Xi'an University of Science and Technology.}
\thanks{S. Zu is with College of Geophysics, Chengdu University of Technology.}
\thanks{The research is supported by the starting fund from Zhejiang University and the National Natural Science Foundation of China (Grant No. 41804140), the Open Fund of Key Laboratory of Exploration Technologies for Oil and Gas Resources (Yangtze University), Ministry of Education (Grant No. PI2018-02). (Corresponding author: Yangkang Chen.)}}
\maketitle

%, college of geology and enviroment
%58 Yanta road
%Beilin district

\begin{abstract}
A sparse dictionary is more adaptive than a sparse fixed-basis transform since it can learn the features directly from the input data in a data-driven way.  However, learning a sparse dictionary is time consuming because a large number of iterations are required in order to obtain the dictionary atoms that best represent the features of input data. The computational cost becomes unaffordable when it comes to high-dimensional problems, e.g., 3D or even 5D applications. We propose an efficient high-dimensional dictionary learning method by avoiding the singular value decomposition (SVD) calculation in each dictionary update step that is required by the classic K-singular value decomposition (KSVD) algorithm. Besides, due to the special structure of the sparse coefficient matrix, it requires a much less expensive sparse coding process. The overall computational advantage of the new dictionary learning method is very decent while the results are still comparable or event better than those from the traditional KSVD method. We apply the proposed method to both 3D and 5D seismic data reconstructions and demonstrate the successful and efficient performance.
\end{abstract}

\begin{keywords}
Seismic data reconstruction, signal processing, dictionary learning, high-dimensional problem
\end{keywords}

@@

\begin{figure}[htb!]
\centering
\subfigure[]{\includegraphics[width=0.3\columnwidth]{synth/Fig/synth-clean}
\label{fig:synth-clean}}
   \subfigure[]{\includegraphics[width=0.3\columnwidth]{synth/Fig/synth-noisy}
   \label{fig:synth-noisy}}
   \subfigure[]{\includegraphics[width=0.3\columnwidth]{synth/Fig/synth-obs}
   \label{fig:synth-obs}}
\caption{Synthetic example with planar events. (a) Clean data. (b) Noisy data (SNR=-0.05 dB). (c) Observed data with 50\% randomly missing traces (SNR=-0.03 dB).}
\label{fig:synth-clean,synth-noisy,synth-obs}
\end{figure}

\begin{figure}[htb!]
\centering
\subfigure[]{\includegraphics[width=0.3\columnwidth]{synth/Fig/synth-ksvd}
\label{fig:synth-ksvd}}
   \subfigure[]{\includegraphics[width=0.3\columnwidth]{synth/Fig/synth-ddtf}
   \label{fig:synth-ddtf}}
   \subfigure[]{\includegraphics[width=0.3\columnwidth]{synth/Fig/synth-sgk}
   \label{fig:synth-sgk}}
\caption{Reconstructed data using (a) KSVD (SNR=8.89 dB), (b) DDTF (SNR=5.42 dB), and (c) SGK (SNR=9.74 dB).}
\label{fig:synth-ksvd,synth-ddtf,synth-sgk}
\end{figure}

\begin{figure}[htb!]
\centering
\subfigure[]{\includegraphics[width=0.45\columnwidth]{Fig/atom_dct}
\label{fig:atom_dct}}
   \subfigure[]{\includegraphics[width=0.45\columnwidth]{Fig/atom_ksvd}
   \label{fig:atom_ksvd}}\\
   \subfigure[]{\includegraphics[width=0.45\columnwidth]{Fig/atom_ddtf}
   \label{fig:atom_ddtf}}
   \subfigure[]{\includegraphics[width=0.45\columnwidth]{Fig/atom_sgk}
   \label{fig:atom_sgk}}
\caption{Comparison of different dictionary atoms. (a) Dictionary atoms of the discrete cosine transform. (b) Dictionary atoms of the KSVD method. (c) Dictionary atoms of the DDTF method.  (d) Dictionary atoms of the SGK method. }
\label{fig:atom_dct,atom_ksvd,atom_ddtf,atom_sgk}
\end{figure}

@@

\begin{figure}[htb!]
\centering
\subfigure[]{\includegraphics[width=0.30\columnwidth]{synth/Fig/synth-ksvd-simi}
\label{fig:synth-ksvd-simi}}
   \subfigure[]{\includegraphics[width=0.30\columnwidth]{synth/Fig/synth-ddtf-simi}
   \label{fig:synth-ddtf-simi}}
   \subfigure[]{\includegraphics[width=0.30\columnwidth]{synth/Fig/synth-sgk-simi}
   \label{fig:synth-sgk-simi}}
\caption{Similarity comparison of reconstructed data using (a) KSVD, (b) DDTF, and (c) SGK.}
\label{fig:synth-ksvd-simi,synth-ddtf-simi,synth-sgk-simi}
\end{figure}

\begin{figure}[htb!]
\centering
	\subfigure[]{\includegraphics[width=0.31\columnwidth]{synth/Fig/synth-s-clean}
	\label{fig:synth-s-clean}}
   \subfigure[]{\includegraphics[width=0.31\columnwidth]{synth/Fig/synth-s-noisy}
   \label{fig:synth-s-noisy}}
   \subfigure[]{\includegraphics[width=0.31\columnwidth]{synth/Fig/synth-s-obs}
   \label{fig:synth-s-obs}}\\
\subfigure[]{\includegraphics[width=0.31\columnwidth]{synth/Fig/synth-s-ksvd}
\label{fig:synth-s-ksvd}}
   \subfigure[]{\includegraphics[width=0.31\columnwidth]{synth/Fig/synth-s-ddtf}
   \label{fig:synth-s-ddtf}}
   \subfigure[]{\includegraphics[width=0.31\columnwidth]{synth/Fig/synth-s-sgk}
   \label{fig:synth-s-sgk}}
\caption{\new{2D slice comparison of the synthetic example. (a) Clean data. (b) Noisy data. (c) Observed data. (d) Reconstructed data using KSVD. (e) Reconstructed data using DDTF. (f) Reconstructed data using SGK.}}
\label{fig:synth-s-clean,synth-s-noisy,synth-s-obs,synth-s-ksvd,synth-s-ddtf,synth-s-sgk}
\end{figure}

\begin{figure}[htb!]
\centering
\subfigure[]{\includegraphics[width=\columnwidth]{synth/Fig/synth-ss}}
\caption{Single-trace comparison. The black line corresponds to the ground-truth solution. The green line corresponds to the noisy data. The pink line corresponds to the proposed method, which is closest to the black line. The red line corresponds to the KSVD method. The blue line corresponds to the DDTF method.}
\label{fig:synth-ss}
\end{figure}

\begin{figure}[htb!]
\centering
\subfigure[]{\includegraphics[width=0.30\columnwidth]{synth/Fig/synth-mssa}
\label{fig:synth-mssa}}
   \subfigure[]{\includegraphics[width=0.30\columnwidth]{synth/Fig/synth-mssasgk}
   \label{fig:synth-mssasgk}}
\caption{Reconstructed data using (a) MSSA (SNR=9.51 dB), and (b) SGK with (a) as the initialization model (SNR=12.22 dB).}
\label{fig:synth-mssa,synth-mssasgk}
\end{figure}

@@

\begin{figure}[htb!]
\centering
\subfigure[]{\includegraphics[width=0.30\columnwidth]{hyper/Fig/hyper-clean}
\label{fig:hyper-clean}}
  \subfigure[]{\includegraphics[width=0.30\columnwidth]{hyper/Fig/hyper-noisy}
  \label{fig:hyper-noisy}}
  \subfigure[]{\includegraphics[width=0.30\columnwidth]{hyper/Fig/hyper-obs}
  \label{fig:hyper-obs}}
\caption{Synthetic example with hyperboloid events. (a) Clean data. (b) Noisy data (SNR=6.93 dB). (c) Observed data with 50\% randomly missing traces (SNR=2.21 dB).}
\label{fig:hyper-clean,hyper-noisy,hyper-obs}
\end{figure}

\begin{figure}[htb!]
\centering
\subfigure[]{\includegraphics[width=0.30\columnwidth]{hyper/Fig/hyper-mssa}
\label{fig:hyper-mssa}}
\subfigure[]{\includegraphics[width=0.30\columnwidth]{hyper/Fig/hyper-ksvd}
\label{fig:hyper-ksvd}}\\
  \subfigure[]{\includegraphics[width=0.30\columnwidth]{hyper/Fig/hyper-ddtf}
  \label{fig:hyper-ddtf}}
  \subfigure[]{\includegraphics[width=0.30\columnwidth]{hyper/Fig/hyper-sgk}
  \label{fig:hyper-sgk}}
\caption{Reconstructed data using (a) Rank-reduction method (SNR=9.68 dB), (b) KSVD (SNR=10.84 dB), (c) DDTF (SNR=11.45 dB), and (d) SGK (SNR=11.00 dB). All DL based methods use the result from the rank-reduction method as the initial model.}
\label{fig:hyper-mssa,hyper-ksvd,hyper-ddtf,hyper-sgk}
\end{figure}

\begin{figure}[htb!]
\centering
\subfigure[]{\includegraphics[width=0.30\columnwidth]{hyper/Fig/hyper-ksvd-simi}
\label{fig:hyper-ksvd-simi}}
  \subfigure[]{\includegraphics[width=0.30\columnwidth]{hyper/Fig/hyper-ddtf-simi}
  \label{fig:hyper-ddtf-simi}}
  \subfigure[]{\includegraphics[width=0.30\columnwidth]{hyper/Fig/hyper-sgk-simi}
  \label{fig:hyper-sgk-simi}}
\caption{Similarity comparison of reconstructed data using (a) KSVD, (b) DDTF, and (c) SGK.}
\label{fig:hyper-ksvd-simi,hyper-ddtf-simi,hyper-sgk-simi}
\end{figure}

\begin{figure}[htb!]
\centering
\subfigure[]{\includegraphics[width=0.31\columnwidth]{hyper/Fig/hyper-s-clean}
\label{fig:hyper-s-clean}}
  \subfigure[]{\includegraphics[width=0.31\columnwidth]{hyper/Fig/hyper-s-noisy}
  \label{fig:hyper-s-noisy}}
  \subfigure[]{\includegraphics[width=0.31\columnwidth]{hyper/Fig/hyper-s-obs}
  \label{fig:hyper-s-obs}}\\
\subfigure[]{\includegraphics[width=0.31\columnwidth]{hyper/Fig/hyper-s-ksvd}
\label{fig:hyper-s-ksvd}}
  \subfigure[]{\includegraphics[width=0.31\columnwidth]{hyper/Fig/hyper-s-ddtf}
  \label{fig:hyper-s-ddtf}}
  \subfigure[]{\includegraphics[width=0.31\columnwidth]{hyper/Fig/hyper-s-sgk}
  \label{fig:hyper-s-sgk}}
\caption{\new{2D slice comparison of the synthetic example. (a) Clean data. (b) Noisy data. (c) Observed data. (d) Reconstructed data using KSVD. (e) Reconstructed data using DDTF. (f) Reconstructed data using SGK.}}
\label{fig:hyper-s-clean,hyper-s-noisy,hyper-s-obs,hyper-s-ksvd,hyper-s-ddtf,hyper-s-sgk}
\end{figure}

@@

\begin{figure}[htb!]
\centering
\subfigure[]{\includegraphics[width=0.31\columnwidth]{synth5d/Fig/syn5d-clean}
\label{fig:syn5d-clean}}
  \subfigure[]{\includegraphics[width=0.31\columnwidth]{synth5d/Fig/syn5d-noisy}
  \label{fig:syn5d-noisy}}
  \subfigure[]{\includegraphics[width=0.31\columnwidth]{synth5d/Fig/syn5d-obs}
  \label{fig:syn5d-obs}}\\
\subfigure[]{\includegraphics[width=0.31\columnwidth]{synth5d/Fig/syn5d-ksvd}
\label{fig:syn5d-ksvd}}
  \subfigure[]{\includegraphics[width=0.31\columnwidth]{synth5d/Fig/syn5d-ddtf}
  \label{fig:syn5d-ddtf}}
  \subfigure[]{\includegraphics[width=0.31\columnwidth]{synth5d/Fig/syn5d-sgk}
  \label{fig:syn5d-sgk}}
\caption{\new{Comparison of 5D seismic reconstruction in a common midpoint gather. (a) Clean data. (b) Noisy data (SNR=-0.65 dB). (c) Decimated data with 70\% randomly removed traces  (SNR=-0.21 dB). (d) Reconstructed data using KSVD  (SNR=6.58 dB). (e) Reconstructed data using DDTF  (SNR=3.22 dB). (f) Reconstructed data using SGK (SNR=6.58 dB).}}
\label{fig:syn5d-clean,syn5d-noisy,syn5d-obs,syn5d-ksvd,syn5d-ddtf,syn5d-sgk}
\end{figure}

\begin{figure}[htb!]
\centering
\subfigure[]{\includegraphics[width=0.31\columnwidth]{synth5d/Fig/syn5d-clean2}
\label{fig:syn5d-clean2}}
  \subfigure[]{\includegraphics[width=0.31\columnwidth]{synth5d/Fig/syn5d-noisy2}
  \label{fig:syn5d-noisy2}}
  \subfigure[]{\includegraphics[width=0.31\columnwidth]{synth5d/Fig/syn5d-obs2}
  \label{fig:syn5d-obs2}}\\
\subfigure[]{\includegraphics[width=0.31\columnwidth]{synth5d/Fig/syn5d-ksvd2}
\label{fig:syn5d-ksvd2}}
  \subfigure[]{\includegraphics[width=0.31\columnwidth]{synth5d/Fig/syn5d-ddtf2}
  \label{fig:syn5d-ddtf2}}
  \subfigure[]{\includegraphics[width=0.31\columnwidth]{synth5d/Fig/syn5d-sgk2}
  \label{fig:syn5d-sgk2}}
\caption{\new{Comparison of 5D seismic reconstruction in a common offset gather. (a) Clean data. (b) Noisy data (SNR=-0.65 dB). (c) Decimated data with 70\% randomly removed traces  (SNR=-0.21 dB). (d) Reconstructed data using KSVD  (SNR=6.58 dB). (e) Reconstructed data using DDTF  (SNR=3.22 dB). (f) Reconstructed data using SGK (SNR=6.58 dB).}}
\label{fig:syn5d-clean2,syn5d-noisy2,syn5d-obs2,syn5d-ksvd2,syn5d-ddtf2,syn5d-sgk2}
\end{figure}

\new{Despite of the highly incomplete sampling of the raw dataset, it is not difficult to learn dictionaries.  In the proposed framework, one can use two options to learn dictionaries from observed data. First, one can directly learn the dictionaries from incomplete data with missing traces. Secondly, one can also learn the dictionaries for initially reconstructed data (e.g., using a rank-based method). Both strategies will be illustrated later. }

\new{An interesting note is that enforcing $\mathbf{s}_j =\mathbf{e}_p$ ($P=1$) helps deriving the SVD-free dictionary updating scheme, as an averaged form of the training samples (e.g., in equation \ref{eq:dk}). $P=1$ is only required in the dictionary learning process. When applied to denoising or reconstruction, the sparse coding process expressed in equation \ref{eq:eq1}, (given the learned dictionary $\mathbf{D}$ from the SGK method), can be taken for the denoising or reconstruction purpose.}

@@

\begin{figure}[htb!]
\centering
\subfigure[]{\includegraphics[width=0.30\columnwidth]{field/Fig/f-obs}
\label{fig:f-obs}}
\subfigure[]{\includegraphics[width=0.30\columnwidth]{field/Fig/f-ksvd}
\label{fig:f-ksvd}}\\
  \subfigure[]{\includegraphics[width=0.30\columnwidth]{field/Fig/f-ddtf}
  \label{fig:f-ddtf}}
  \subfigure[]{\includegraphics[width=0.30\columnwidth]{field/Fig/f-sgk}
  \label{fig:f-sgk}}
\caption{\new{3D Field data example. (a) Observed data. Reconstructed data using (b) KSVD, (c) DDTF, and (d) SGK.} }
\label{fig:f-obs,f-ksvd,f-ddtf,f-sgk}
\end{figure}

\begin{figure}[htb!]
\centering
\subfigure[]{\includegraphics[width=0.30\columnwidth]{field/Fig/f-s-obs0}
\label{fig:f-s-obs0}}
\subfigure[]{\includegraphics[width=0.30\columnwidth]{field/Fig/f-s-ksvd0}
\label{fig:f-s-ksvd0}}\\
  \subfigure[]{\includegraphics[width=0.30\columnwidth]{field/Fig/f-s-ddtf0}
  \label{fig:f-s-ddtf0}}
  \subfigure[]{\includegraphics[width=0.30\columnwidth]{field/Fig/f-s-sgk0}
  \label{fig:f-s-sgk0}}
\caption{2D slice comparison of the field data example. (a) Observed data. (b) Reconstructed data using KSVD. (c) Reconstructed data using DDTF. (d) Reconstructed data using SGK. The frame boxes are zoomed and shown in Figure \ref{fig:fz-obs,fz-ksvd,fz-ddtf,fz-sgk}. }
\label{fig:f-s-obs0,f-s-ksvd0,f-s-ddtf0,f-s-sgk0}
\end{figure}

\begin{figure}[htb!]
\centering
\subfigure[]{\includegraphics[width=0.30\columnwidth]{field/Fig/fz-obs}
\label{fig:fz-obs}}
\subfigure[]{\includegraphics[width=0.30\columnwidth]{field/Fig/fz-ksvd}
\label{fig:fz-ksvd}}\\
  \subfigure[]{\includegraphics[width=0.30\columnwidth]{field/Fig/fz-ddtf}
  \label{fig:fz-ddtf}}
  \subfigure[]{\includegraphics[width=0.30\columnwidth]{field/Fig/fz-sgk}
  \label{fig:fz-sgk}}
\caption{Zoomed 2D slice comparison of the field data example. (a) Observed data. (b) Reconstructed data using KSVD. (c) Reconstructed data using DDTF. (d) Reconstructed data using SGK. }
\label{fig:fz-obs,fz-ksvd,fz-ddtf,fz-sgk}
\end{figure}

\begin{figure}[htb!]
\centering
\subfigure[]{\includegraphics[width=0.45\columnwidth]{field5d/Fig/f5d-obs}
\label{fig:f5d-obs}}
\subfigure[]{\includegraphics[width=0.45\columnwidth]{field5d/Fig/f5d-ksvd}
\label{fig:f5d-ksvd}}\\
  \subfigure[]{\includegraphics[width=0.45\columnwidth]{field5d/Fig/f5d-ddtf}
  \label{fig:f5d-ddtf}}
  \subfigure[]{\includegraphics[width=0.45\columnwidth]{field5d/Fig/f5d-sgk}
  \label{fig:f5d-sgk}}
\caption{5D seismic reconstruction example. A common offset gather is displayed for comparison. The 3D common offset gather has been reformulated as 2D matrix. (a) Observed data with approximately 70\% missing traces. (b) Reconstructed data using KSVD. (c) Reconstructed data using DDTF. (d) Reconstructed data using SGK. }
\label{fig:f5d-obs,f5d-ksvd,f5d-ddtf,f5d-sgk}
\end{figure}

\begin{figure}[ht!]
\centering
\subfigure[]{\includegraphics[width=0.45\columnwidth]{field5d/Fig/z-obs}
\label{fig:z-obs}}
\subfigure[]{\includegraphics[width=0.45\columnwidth]{field5d/Fig/z-ksvd}
\label{fig:z-ksvd}}\\
  \subfigure[]{\includegraphics[width=0.45\columnwidth]{field5d/Fig/z-ddtf}
  \label{fig:z-ddtf}}
  \subfigure[]{\includegraphics[width=0.45\columnwidth]{field5d/Fig/z-sgk}
  \label{fig:z-sgk}}
\caption{Zoomed comparison of the 5D seismic reconstruction example. The zooming area is stressed by the frame boxes in Figure \ref{fig:f5d-obs,f5d-ksvd,f5d-ddtf,f5d-sgk}. (a) Observed data with approximately 70\% missing traces. (b) Reconstructed data using KSVD. (c) Reconstructed data using DDTF. (d) Reconstructed data using SGK. The SGK result is obviously better than the other two methods.}
\label{fig:z-obs,z-ksvd,z-ddtf,z-sgk}
\end{figure}

@@

>>>>

\published{IEEE Transactions on Geoscience and Remote Sensing, xxx, xxx, (2021)}

\title{Fast Dictionary Learning for High-Dimensional Seismic Reconstruction}
\author{Hang Wang\footnotemark[1], Wei Chen\footnotemark[2]\footnotemark[3], Quan Zhang\footnotemark[1], Xingye Liu\footnotemark[4], Shaohuan Zu\footnotemark[5], and Yangkang Chen\footnotemark[1]}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\ms{TGRS-2021}

\address{
\footnotemark[1]
School of Earth Sciences\\
Zhejiang University\\
Hangzhou, Zhejiang Province, China, 310027\\
%908352517@qq.com (hang)
%yangkang.chen@zju.edu.cn \\
\footnotemark[2]
Key Laboratory of Exploration Technology for Oil and Gas Resources\\
Ministry of Education\\
Yangtze University\\
\footnotemark[3]
Hubei Cooperative Innovation Center of Unconventional Oil and Gas\\
Yangtze University \\
\footnotemark[4]College of Geology and Environment\\
Shaanxi Provincial Key Laboratory of Geological Support for Coal Green Exploitation\\
Xi'an University of Science and Technology\\
\footnotemark[5] College of Geophysics\\
Chengdu University of Technology
}

\lefthead{Wang et al., 2021}
\righthead{SGK5D}

\maketitle

\begin{abstract}
A sparse dictionary is more adaptive than a sparse fixed-basis transform since it can learn the features directly from the input data in a data-driven way.  However, learning a sparse dictionary is time consuming because a large number of iterations are required in order to obtain the dictionary atoms that best represent the features of input data. The computational cost becomes unaffordable when it comes to high-dimensional problems, e.g., 3D or even 5D applications. We propose an efficient high-dimensional dictionary learning method by avoiding the singular value decomposition (SVD) calculation in each dictionary update step that is required by the classic K-singular value decomposition (KSVD) algorithm. Besides, due to the special structure of the sparse coefficient matrix, it requires a much less expensive sparse coding process. The overall computational advantage of the new dictionary learning method is very decent while the results are still comparable or event better than those from the traditional KSVD method. We apply the proposed method to both 3D and 5D seismic data reconstructions and demonstrate the successful and efficient performance.\\
\textbf{Keywords}\\
Seismic data reconstruction, signal processing, dictionary learning, high-dimensional problem
\end{abstract}

@@

\inputdir{synth}
\multiplot{3}{synth-clean,synth-noisy,synth-obs}{width=0.3\columnwidth}{Synthetic example with planar events. (a) Clean data. (b) Noisy data (SNR=-0.05 dB). (c) Observed data with 50\% randomly missing traces (SNR=-0.03 dB).}

\multiplot{3}{synth-ksvd,synth-ddtf,synth-sgk}{width=0.3\columnwidth}{Reconstructed data using (a) KSVD (SNR=8.89 dB), (b) DDTF (SNR=5.42 dB), and (c) SGK (SNR=9.74 dB).}

\inputdir{./}
\multiplot{4}{atom_dct,atom_ksvd,atom_ddtf,atom_sgk}{width=0.45\columnwidth}{Comparison of different dictionary atoms. (a) Dictionary atoms of the discrete cosine transform. (b) Dictionary atoms of the KSVD method. (c) Dictionary atoms of the DDTF method.  (d) Dictionary atoms of the SGK method.}

@@

\inputdir{synth}
\multiplot{3}{synth-ksvd-simi,synth-ddtf-simi,synth-sgk-simi}{width=0.3\columnwidth}{Similarity comparison of reconstructed data using (a) KSVD, (b) DDTF, and (c) SGK.}

\multiplot{6}{synth-s-clean,synth-s-noisy,synth-s-obs,synth-s-ksvd,synth-s-ddtf,synth-s-sgk}{width=0.3\columnwidth}{\new{2D slice comparison of the synthetic example. (a) Clean data. (b) Noisy data. (c) Observed data. (d) Reconstructed data using KSVD. (e) Reconstructed data using DDTF. (f) Reconstructed data using SGK.}}

\plot{synth-ss}{width=\columnwidth}{width=0.3\columnwidth}{Single-trace comparison. The black line corresponds to the ground-truth solution. The green line corresponds to the noisy data. The pink line corresponds to the proposed method, which is closest to the black line. The red line corresponds to the KSVD method. The blue line corresponds to the DDTF method.}

\multiplot{2}{synth-mssa,synth-mssasgk}{width=0.30\columnwidth}{Reconstructed data using (a) MSSA (SNR=9.51 dB), and (b) SGK with (a) as the initialization model (SNR=12.22 dB).}

@@

\inputdir{hyper}
\multiplot{3}{hyper-clean,hyper-noisy,hyper-obs}{width=0.30\columnwidth}{Synthetic example with hyperboloid events. (a) Clean data. (b) Noisy data (SNR=6.93 dB). (c) Observed data with 50\% randomly missing traces (SNR=2.21 dB).}

\multiplot{4}{hyper-mssa,hyper-ksvd,hyper-ddtf,hyper-sgk}{width=0.30\columnwidth}{Reconstructed data using (a) Rank-reduction method (SNR=9.68 dB), (b) KSVD (SNR=10.84 dB), (c) DDTF (SNR=11.45 dB), and (d) SGK (SNR=11.00 dB). All DL based methods use the result from the rank-reduction method as the initial model.}

\multiplot{3}{hyper-ksvd-simi,hyper-ddtf-simi,hyper-sgk-simi}{width=0.30\columnwidth}{Similarity comparison of reconstructed data using (a) KSVD, (b) DDTF, and (c) SGK.}

\multiplot{6}{hyper-s-clean,hyper-s-noisy,hyper-s-obs,hyper-s-ksvd,hyper-s-ddtf,hyper-s-sgk}{width=0.30\columnwidth}{\new{2D slice comparison of the synthetic example. (a) Clean data. (b) Noisy data. (c) Observed data. (d) Reconstructed data using KSVD. (e) Reconstructed data using DDTF. (f) Reconstructed data using SGK.}}

\inputdir{synth5d}
\multiplot{6}{syn5d-clean,syn5d-noisy,syn5d-obs,syn5d-ksvd,syn5d-ddtf,syn5d-sgk}{width=0.30\columnwidth}{\new{Comparison of 5D seismic reconstruction in a common midpoint gather. (a) Clean data. (b) Noisy data (SNR=-0.65 dB). (c) Decimated data with 70\% randomly removed traces  (SNR=-0.21 dB). (d) Reconstructed data using KSVD  (SNR=6.58 dB). (e) Reconstructed data using DDTF  (SNR=3.22 dB). (f) Reconstructed data using SGK (SNR=6.58 dB).}}
\multiplot{6}{syn5d-clean2,syn5d-noisy2,syn5d-obs2,syn5d-ksvd2,syn5d-ddtf2,syn5d-sgk2}{width=0.30\columnwidth}{\new{Comparison of 5D seismic reconstruction in a common offset gather. (a) Clean data. (b) Noisy data (SNR=-0.65 dB). (c) Decimated data with 70\% randomly removed traces  (SNR=-0.21 dB). (d) Reconstructed data using KSVD  (SNR=6.58 dB). (e) Reconstructed data using DDTF  (SNR=3.22 dB). (f) Reconstructed data using SGK (SNR=6.58 dB).}}

@@

Despite of the highly incomplete sampling of the raw dataset, it is not difficult to learn dictionaries.  In the proposed framework, one can use two options to learn dictionaries from observed data. First, one can directly learn the dictionaries from incomplete data with missing traces. Secondly, one can also learn the dictionaries for initially reconstructed data (e.g., using a rank-based method). Both strategies will be illustrated later.

An interesting note is that enforcing $\mathbf{s}_j =\mathbf{e}_p$ ($P=1$) helps deriving the SVD-free dictionary updating scheme, as an averaged form of the training samples (e.g., in equation \ref{eq:dk}). $P=1$ is only required in the dictionary learning process. When applied to denoising or reconstruction, the sparse coding process expressed in equation \ref{eq:eq1}, (given the learned dictionary $\mathbf{D}$ from the SGK method), can be taken for the denoising or reconstruction purpose.

@@

\inputdir{field}
\multiplot{4}{f-obs,f-ksvd,f-ddtf,f-sgk}}{width=0.30\columnwidth}{\new{3D Field data example. (a) Observed data. Reconstructed data using (b) KSVD, (c) DDTF, and (d) SGK.}}
\multiplot{4}{f-s-obs0,f-s-ksvd0,f-s-ddtf0,f-s-sgk0}{width=0.30\columnwidth}{2D slice comparison of the field data example. (a) Observed data. (b) Reconstructed data using KSVD. (c) Reconstructed data using DDTF. (d) Reconstructed data using SGK. The frame boxes are zoomed and shown in Figure \ref{fig:fz-obs,fz-ksvd,fz-ddtf,fz-sgk}.}

\multiplot{4}{fz-obs,fz-ksvd,fz-ddtf,fz-sgk}{width=0.30\columnwidth}{Zoomed 2D slice comparison of the field data example. (a) Observed data. (b) Reconstructed data using KSVD. (c) Reconstructed data using DDTF. (d) Reconstructed data using SGK. }

\inputdir{./}
\multiplot{4}{f5d-obs,f5d-ksvd,f5d-ddtf,f5d-sgk}{width=0.45\columnwidth}{5D seismic reconstruction example. A common offset gather is displayed for comparison. The 3D common offset gather has been reformulated as 2D matrix. (a) Observed data with approximately 70\% missing traces. (b) Reconstructed data using KSVD. (c) Reconstructed data using DDTF. (d) Reconstructed data using SGK. }

\multiplot{4}{z-obs,z-ksvd,z-ddtf,z-sgk}{width=0.45\columnwidth}{Zoomed comparison of the 5D seismic reconstruction example. The zooming area is stressed by the frame boxes in Figure \ref{fig:f5d-obs,f5d-ksvd,f5d-ddtf,f5d-sgk}. (a) Observed data with approximately 70\% missing traces. (b) Reconstructed data using KSVD. (c) Reconstructed data using DDTF. (d) Reconstructed data using SGK. The SGK result is obviously better than the other two methods. }

@@

\section{Acknowledgement}
The research is supported by the starting fund from Zhejiang University and the National Natural Science Foundation of China (Grant No. 41804140), the Open Fund of Key Laboratory of Exploration Technologies for Oil and Gas Resources (Yangtze University), Ministry of Education (Grant No. PI2018-02).
